---
title: "Rworksheet#4b_Caballero"
author: "Jireh Niel Caballero"
date: "2023-11-22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#1
vectorA <- c(1,2,3,4,5)
vectorA

mat <- matrix(0, nrow=5, ncol=5)
mat

#     [,1] [,2] [,3] [,4] [,5]
#[1,]    0    0    0    0    0
#[2,]    0    0    0    0    0
#[3,]    0    0    0    0    0
#[4,]    0    0    0    0    0
#[5,]    0    0    0    0    0


for(i in 1:5){
  for(j in 1:5){
      mat[i,j] <- abs(vectorA[i] - vectorA[j])
    }
  }
print(mat)

#     [,1] [,2] [,3] [,4] [,5]
#[1,]    0    1    2    3    4
#[2,]    1    0    1    2    3
#[3,]    2    1    0    1    2
#[4,]    3    2    1    0    1
#[5,]    4    3    2    1    0
```

```{r}
#2
rightTriangle <- c()

for (i in 1:5){
  for(j in 1:i+1){
    rightTriangle = c(rightTriangle,"*")
  }
  print(rightTriangle)
  rightTriangle <- c()
}

#[1] "*"
#[1] "*" "*"
#[1] "*" "*" "*"
#[1] "*" "*" "*" "*"
#[1] "*" "*" "*" "*" "*"
```

```{r}
#3
n <- as.integer(readline(prompt = "Enter the number of terms: "))

#n <- as.integer(readline(prompt = "Enter the number of terms: "))
#Enter the number of terms: 

a<- 0
b<- 1

cat("Fibonacci sequnce: ",a,b)
#Fibonacci sequnce:  0 1

repeat{
  c<-a+b
  if (c>500){
    break
  }
  cat(",",c)
  a<-b
  b<-c
}
#, 1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 144, 233, 377
```


```{r}
#using basic Graphics(plot(),barplot(),pie(),hist()) 


#4

#a
library(readr)
library(openxlsx)

HouseholdData <- read.table("/cloud/project/CaballeroRworksheet#4/FinalHouseholdData.csv", header = TRUE, sep = ",", as.is = TRUE)
#View(HouseholdData)
#head(HouseholdData,6)
#  X shoe_size Height  Gender
#1 1       6.5   66.0   F
#2 2       9.0   68.0   F
#3 3       8.5   64.5   F
#4 4       8.5   65.0   F
#5 5      10.5   70.0   M
#6 6       7.0   64.0   F
```


#b

```{r}
maleSubset <- subset(HouseholdData, HouseholdData$Gender == 'M')
femaleSubset <- subset(HouseholdData, HouseholdData$Gender == 'F')


male_count <- nrow(maleSubset)
#[1] 14
female_count <- nrow(femaleSubset)
#[1] 14

#C
```



```{r}
count <- c(male_count, female_count)
gender <- c("Male", "Female")

barplot(count,
        names.arg = gender,
        main = "The number of Males and Females in Household Data",
        xlab = "gender",
        ylab = "Count",
        col = c("blue", "pink"),
        border = "black")

# Add legend
legend("topright",
       legend = gender,
       fill = c("blue", "pink"))

```

```{r}
#5

#a
monthly_income<- c(60,10,5,25)

month_labels <- round(monthly_income/sum(monthly_income)*100,1)
month_labels <- paste(month_labels,"%", sep = "")

pie(monthly_income, main = "The monthly income of Dela Cruz family",
    col = rainbow(length(monthly_income)),
    labels = month_labels,
    cex = 0.8)

legend(1.5,0.5,
       c("Food","Electricity","Savings","Miscellaneous"),
       fill = rainbow(length(monthly_income)))



```

```{r}
data(iris)
str(iris)
```
#the str(iris)output give us the dataset of iris


```{r}
mean<-colMeans(iris[, c("Sepal.Length","Petal.Length", "Petal.Width")])
mean
```
#c

```{r}
pie(table(iris$Species),
    main = "Species distribution",
    labels = levels(iris$Species),
    col = c("lightblue","lightgreen","red"))

legend("topright", legend = levels(iris$Species),
       fill = c("lightblue","lightgreen","red"),
       title = "Setosa", "Versicolor", "Virginica")
```
#d
```{r}
setosa_lastsix<- tail(subset(iris,Species == "setosa"), n = 6)
versicolor_lastsix<- tail(subset(iris,Species == "versicolor"), n =6)
virginica_lastsix<- tail(subset(iris,Species == "virginica"), n = 6)
setosa_lastsix
```

```{r}
versicolor_lastsix
```
```{r}
virginica_lastsix
```

#E
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
pch = as.integer(iris$Species),
col = as.integer(iris$Species) + 1,
main = "Iris Dataset",
sub = "Sepal width and length",
xlab = "Sepal Length",
ylab = "Sepal Width",
cex = 1.5,
lwd = 1.5)

```

```{r}
as.factor(iris$Species)
```

##as.factor(iris$Species) is a way of telling R that the "Species" variable should be treated as a categorical variable. 



##Basic Cleaning and Transformation of Objects

#7

#A
```{r}
library(readr)
library(readxl)
alexa_file <- read_excel("/cloud/project/CaballeroRworksheet4b/alexa_file.xlsx")
#View(alexa_file)

```

```{r}
alexaVaration <- gsub("Black Plus", "Black Plus", alexa_file$variation)
alexa_file$variation <- gsub("Black Show", "Black Show", alexa_file$variation)
alexa_file$variation <- gsub("Black Spot", "Black Spot", alexa_file$variation)
alexa_file$variation <- gsub("Black Dot", "Black Dot", alexa_file$variation)
alexa_file$variation <- gsub("White Dot", "White Dot", alexa_file$variation)
alexa_file$variation <- gsub("White Plus", "White Plus", alexa_file$variation)
alexa_file$variation <- gsub("White Show", "White Show", alexa_file$variation)
alexa_file$variation <- gsub("White Spot", "White Spot", alexa_file$variation)



```

#B
```{r}
library(dplyr)
save(alexa_file, file = "variations.RData")
load("variations.RData")
alexaVaration <- alexa_file%>%count(alexa_file$variation)
alexaVaration
```
#alexa_file$variation
#<chr>
#n
#<int>
#Black	261			
#Black Dot	516			
#Black Plus	270			
#Black Show	265			
#Black Spot	241			
#Charcoal Fabric	430			
#Configuration: Fire TV Stick	350			
#Heather Gray Fabric	157			
#Oak Finish	14			
#Sandstone Fabric	90	
#Walnut Finish	9			
#White	91			
#White Dot	184			
#White Plus	78			
#White Show	85			
#White Spot	109	

#C
```{r}
barplot(
height = alexaVaration$n,
names.arg = alexaVaration$`alexa_file$variation`,
col = "pink",
main = "Alexa Varations",
las = 2,
cex.names = 0.58
)

```
#D

```{r}
par(mfrow = c(1, 2))
black_variants <- alexaVaration[1:5,]
white_variants <- alexaVaration[12:16,]
barplot(
height = black_variants$n,
names.arg = black_variants$`alexa_file$variation`,
main = "Black Variants",
col = rainbow(8),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)
barplot(
height = white_variants$n,
names.arg = white_variants$`alexa_file$variation`,
main = "White Variants",
col = rainbow(8),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)
```



